Supplementary Materials

### 2023-06-08 manuscript/manuscript.tex
new figs

@@

@@

\subsection*{DFE Estimates}
% - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA

>>>>

\subsection*{Annotation Model Comparison}

Our method takes tracks of annotated features that are \emph{a priori} expected
to have similar fitness effects, and estimates the distribution of fitness
effects for each feature type. Additionally, our method estimates the
per-basepair per-generation deleterious mutation rate ($\mu_\text{del}$) and
the background level of pairwise diversity ($\pi_0$). Since the theory used by
our method can in principle accommodate weak selection and neutrality. To test
this, for each our annotation models, we fit (1) ``sparse" models that only
cover the fraction of the genome that is the union of all tracks, and (2)
``full`` models that are sparse models with the remaining complement of the
genome labeled as ``other".

Overall, we find that

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_2.pdf} 
    \caption{The $R^2$ estimates for a sparse (A) and full (B) models, for all 
    populations (colors). Round points are our B' method; diamonds are the 
    classic B. For our method, we also calculate leave-one-out out-sample $R^2$
    (lighter filled points).}
  \label{fig:figure-2}
\end{figure}

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_3.pdf} 
    \caption{The distribution of fitness effects of new mutations estimates for
        Yoruba individuals. (A) The DFEs using sparse (left column) and
        full-coverage (right column) tracks, across different annotation models
        (row). Color indicates the feature type. (B) The DFE of the
        full-coverage CDS-priority model comparing the estimates across
    populations.}
  \label{fig:figure-3}
\end{figure}

@@

\subsection*{DFE Estimates}

\begin{table}
\centering
\caption{$R^2$ and mutation rate estimates for all models}
\begin{tabular}{lll|crr|cc}
    \textbf{Model} & \textbf{Track type} & \textbf{Pop.} & \textbf{B' $R_\text{LOO}^2$} & \textbf{B' $R_\text{IS}^2$} & \textbf{B $R_\text{IS}^2$} & \textbf{B' $\hat{\mu} \times 10^{-8}$} & \textbf{B $\hat{\mu} \times 10^{-8}$} \\[0.5ex] 
\hline
\hline
    \textbf{phastcons$>$CDS$>$genes} &            sparse &          YRI &                        \textbf{68.45} &             68.16 &            65.23 &                                 1.8636 &                                0.2975 \\
phastcons$>$CDS$>$genes &              full &          YRI &                        68.19 &             68.12 &            63.67 &                                 1.7082 &                                0.1666 \\
               CADD 6\% &              full &          YRI &                        68.08 &             68.08 &            62.96 &                                 1.4612 &                                0.1721 \\
               CADD 6\% &            sparse &          YRI &                        68.04 &             68.03 &            68.01 &                                 2.0912 &                                2.0716 \\
               CADD 8\% &              full &          YRI &                        66.98 &             67.11 &            63.61 &                                 1.0126 &                                0.1713 \\
               CADD 8\% &            sparse &          YRI &                        66.67 &             67.01 &            66.98 &                                 1.5688 &                                1.5473 \\
CDS$>$genes$>$phastcons &              full &          YRI &                        64.90 &             65.51 &            61.23 &                                 4.1951 &                                0.1620 \\
CDS$>$genes$>$phastcons &            sparse &          YRI &                        64.90 &             65.51 &            63.16 &                                 4.1795 &                                0.3070 \\
\textbf{phastcons$>$CDS$>$genes} &            sparse &          CEU &                        \textbf{61.98} &             63.02 &            60.02 &                                 2.0194 &                                0.3036 \\
phastcons$>$CDS$>$genes &              full &          CEU &                        61.78 &             63.01 &            58.57 &                                 2.3499 &                                0.1679 \\
               CADD 6\% &              full &          CEU &                        61.74 &             62.94 &            57.42 &                                 1.5365 &                                0.1739 \\
               CADD 6\% &            sparse &          CEU &                        61.66 &             62.86 &            62.84 &                                 2.1215 &                                2.1084 \\
               CADD 8\% &              full &          CEU &                        60.80 &             62.12 &            57.95 &                                 1.1572 &                                0.1742 \\
               CADD 8\% &            sparse &          CEU &                        60.59 &             62.06 &            62.04 &                                 1.6030 &                                1.5876 \\
               CADD 6\% &              full &          CHB &                        59.62 &             61.31 &            55.31 &                                 1.5083 &                                0.1727 \\
               CADD 6\% &            sparse &          CHB &                        59.54 &             61.21 &            61.20 &                                 2.1242 &                                2.1170 \\
               \textbf{phastcons$>$CDS$>$genes} &            sparse &          CHB &                        \textbf{59.44} &             61.08 &            57.88 &                                 2.1982 &                                0.3060 \\
phastcons$>$CDS$>$genes &              full &          CHB &                        59.29 &             61.08 &            56.71 &                                 2.5573 &                                0.1694 \\
               CADD 8\% &              full &          CHB &                        58.88 &             60.63 &            55.86 &                                 1.1128 &                                0.1733 \\
               CADD 8\% &            sparse &          CHB &                        58.65 &             60.57 &            60.55 &                                 1.6098 &                                1.6001 \\
CDS$>$genes$>$phastcons &            sparse &          CEU &                        58.58 &             60.30 &            58.41 &                                 4.1041 &                                0.3110 \\
CDS$>$genes$>$phastcons &              full &          CEU &                        58.58 &             60.30 &            56.75 &                                 4.1001 &                                0.1648 \\
CDS$>$genes$>$phastcons &            sparse &          CHB &                        56.48 &             58.73 &            56.60 &                                 3.7195 &                                0.3136 \\
CDS$>$genes$>$phastcons &              full &          CHB &                        56.48 &             58.73 &            54.95 &                                 3.7552 &                                0.1654 \\
\hline
\end{tabular}
\end{table}

Our best-fitting models have out-sample $R^2$ of around 68\% for Yoruba, 62\%
European, and 60\% for Han Chinese ancestry individuals. However, $R^2$ has a
theoretic upper bound due to coalescence noise around the predicted mean level
$\widehat{\pi}(x)$.

% - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA

### 2023-06-08 manuscript/supplementary.tex
new supp, ms stuff
new figs

>>>>

\section*{Additional Results}

\subsection*{Model Comparisons}

% \begin{table}[ht]
%     \caption{CADD 6\% Sparse Track B and B' Comparison}
%     \label{supptab:cadd6}
%     \centering
% \begin{tabular}{lll}
%                          & \strong{B}              & \strong{B'}                     \\
%                          \hline
%     \emph{negative log-likelihood} & 250,622,462,847.73962     & 250,622,458,048.0578      \\
%     $R^2$                   & 53.89\%                 & 53.9\%                  \\
%     \hline
%     $\pi_0$                 & 0.0014272  & 0.0014297  \\
%     $\mu_\text{del}$        & $1.88 \times 10^{-8}$ & $4.29 \times 10^{-8}$ \\
%     \\
%     \emph{sel. coef.} &\multicolumn{1}{c}{\emph{DFE estimates (percent)}} \\
%     \hline
%     $10^{-8}$                  & 0.0                    & 0.0                    \\
%     $10^{-7}$                   & 0.0                    & 0.0                    \\
%     $10^{-6}$                   & 0.0                    & 0.0                    \\
%     $10^{-5}$                   & 0.3                    & 0.57                   \\
%     $10^{-4}$                  & 10.27                  & 0.02                   \\
%     $10^{-3}$                   & 25.22                  & 0.12                   \\
%     $10^{-2}$                    & 1.91                   & 0.0                    \\
%     $10^{-1}$                     & 62.3                   & 0.28                  \\
%     \hline
% \end{tabular}
% \end{table}

\section*{Approximate $R^2$ Limit}

We derive a rough approximation of the theoretic upper bound of $R^2$ by
noting that,

\begin{align}
    R^2 &= \frac{SS_\text{reg}}{SS_\text{reg} + SS_\text{res}} \\ 
        &= \frac{\sum_i (\pi_i - \widehat{\pi}_i)^2}{\sum_i (\pi_i - \widehat{\pi}_i)^2 + SS_\text{res}}
\end{align}

where $\pi_i$ is the observed diversity in window $i$, and $\widehat{\pi}_i$ is
the predicted diversity under a particular model.

Then, $SS_\text{res}$ is the residual noise due to the randomness in mutation
and average coalescence times in the window. If we make the assumptions that
mutation rates are heterogeneous across the genome and the coalescence rate in
a window is simply a background rate rescaled by the reduction factor $B_i$,
then,

\begin{align}
    SS_\text{res} &= \frac{1}{n} \sum_{i=1}^n \var(w\pi_i) \\ 
                  &= \frac{1}{n} \sum_{i=1}^n \var(w\pi_0 B_i) 
\end{align}

where $w$ is the width of the window such that the per-basepair $\pi$ is
appropriately scaled. Then, we note that the variance in pairwise diversity is
equivalent to the variance in the number of segregating sites for a sample of 
two, $\var(\pi) = \var(S_2)$. From \textcite[eq. 7.20]{Wakeley2009-ua}

\begin{align}
    \var(S_2) = \theta + \theta^2 \frac{2}{\rho^2} \int_0^\rho (\rho - x) f_2(x) dx
\end{align}
where,
\begin{align}
    f_2(\rho) = \frac{\rho + 18}{\rho^2 + 13 \rho + 18}.
\end{align}

Since this is a rough approximation, we set $\theta = w \pi_0 B_i$ and fix
$\rho = w \gamma \pi_0 B_i$ where $\gamma$ is the ratio of recombination to
mutation rates, $\nicefrac{r}{\mu}$. In humans, $\gamma \approx 1$, but we
wanted to check the sensitivity of this calculation to this.

This back-of-the-envelope calculation finds that the upper bound of $R^2$ would
be around 74\% for Yoruba, and 80-81\% for European and Han Chinese models
(Figure \ref{suppfig:r2_upper} We note, however, that this rough approximation
assumes constant demography, and that the out-of-Africa bottleneck causes a
reduction in effective population size that would shrink the variance of
exponential coalescence times, but a bottleneck could in fact inflate these
variances. A more accurate approximate of the upper $R^2$ bound could rely on
simulations with more realistic demography.

\begin{figure}[htbp] \centering
    \includegraphics[width=0.5\textwidth]{figures/supplementary/suppfigure_R2_upper.pdf}
    \caption{A rough approximation to the theoretic $R^2$ upper bound assuming constant demography.}
  \label{suppfig:r2_upper}
\end{figure}
